---
title: "BIOSTAT650 HW7"
author: "Yao Song"
date: "12/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("sas7bdat")
myData = read.sas7bdat("/Users/selina/Desktop/BIOS\ 650/Homework/3/completedata.sas7bdat")
head(myData)
#
```

## Problem 1 

## (a)
Fit a model using fatalism as the outcome (Y) and predictors: age (continuous, linear), sex, and race/ethnicity, number of comorbidities (comorbidity1, continuous, linear) and education.
```{r}
variables <- c("Fatalism", "Age", "Sex", "R_E", "Comorbidity1", "Education")
newData = myData[, variables]
fit_1 = lm(Fatalism ~ Age + Sex + R_E + Comorbidity1 + Education, data = newData)
summary(fit_1)
```


Obtain all residuals:

(1). residuals

```{r}
# Method 1: directly oobtain 
fit_1.res = fit_1$residuals  
summary(fit_1.res)

# Method 2: ei = Fatalism - fit_1.yhat
fit_1.yhat = fit_1$fitted.values
fit_1.res2 = newData$Fatalism - fit_1.yhat
# check equvalence
all(round(fit_1.res, 6) == round(fit_1.res2, 6))
```

(2). standardized residuals 

```{r}
# From regression output, MSE = sigma^2hat = 5.171
n = nrow(newData)
p = fit_1$rank
fit_1.sigma = sqrt(sum(fit_1.res^2)/(n-p))
fit_1.zi = fit_1.res/fit_1.sigma  # standardized residuals
```
```{r}
summary(fit_1.zi) # Saw approximately a Normal(0,1) distribution
```

(3). studentized residuals

```{r}
fit_1.ri = rstandard(fit_1)    # internally studentized residuals
summary(fit_1.ri)
```

(4). externally studentized residuals

```{r}
fit_1.r_i = rstudent(fit_1)    # externally studentized residuals 
summary(fit_1.r_i)
```


## (b)
Visualize the differences among various forms of standardizing residuals.

(1). Create a plot of the difference between studentized and standardized residuals (y-axis) vs. the standardized residuals. Make the size of the dots on the plot depend on the leverage value for the observation. 
```{r}
fit_1.hii = hatvalues(fit_1) # leverage hii

y = fit_1.ri - fit_1.zi
x = Standardized = fit_1.zi

data.frame(y, x, id = 1:nrow(newData))
plot(x, y,
     ylab = "Difference between Studentized and Standardized Residuals", 
     xlab = "Standardized Residuals",
     pch = 21, col = 25, bty = "o", cex = fit_1.hii*80)

text(x[which(y == max(y))], y[which(y == max(y))],
     labels = id[which(y == max(y))],
     cex = 1, pos = 4, col = "red")
```
Discuss which observations fall further from 0 (i.e., which observations have the largest differences in the calculated residuals) in terms of hii, and why you could expect that to be the case.\\



???



(2). Create a plot of the difference between externally studentized residuals and studentized (y-axis) vs. the studentized residuals. Make the size of the dots on the plot depend on the square of the regular residual for the observation (i.e. $\epsilon_{i}^2$). 

```{r}
data.frame(Diff_Studentized_Standardized = fit_1.r_i - fit_1.ri, Standardized = fit_1.ri)
plot(fit_1.ri, fit_1.r_i - fit_1.ri, 
     ylab = "Difference between Externally Studentized and Studentized Residuals", 
     xlab = "Studentized Residuals",
     pch = 21, col = 25, bty = "o", cex = (fit_1.res^2)/80)
```
Discuss which observations have the largest differences in the two residuals in terms of $\epsilon_{i}^2$ difference between externally studentized residuals and studentized:\\


```{r}

```

```{r}

```
```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


